Correlation

### 2022-04-07 graphs_newt_snakes/Correlation.Rmd
I finally got all of my graphs made and writen about for the correlation experiment. That took awhile to do

```{r, Cor data gather}
#Reading in the names of files to cycle through them
lit_5000_files <- list.files(path="~/Desktop/cor_5000/cor_byUnit/text_files", pattern="GA5_lit_", full.names=TRUE, recursive=FALSE)
cor_5000_files <- list.files(path="~/Desktop/cor_5000/cor_byUnit/text_files", pattern="GA5_cor_", full.names=TRUE, recursive=FALSE)

@@

```{r, reusable figures, results='hide', message=FALSE, warning=FALSE}

@@

@@

```{r, lest see that correlation}

@@

```{r, cor 5000 zooom}

@@

```{r, correlation 20000}

@@

```{r, correlation 100000}

@@

```{r, boxplotting 5000}
  print("CA = mean_newt_pheno_By_mean_snake_pheno")
  print("CB = num_newts_By_num_snakes")
  print("CC = sum_newt_pheno_By_num_snake")
  print("CD = sum_snake_pheno_By_num_newt")

@@

print("Difference in simulation length")
grid.arrange(G1, G2, G3, nrow=2)

@@

```{r, correlation 5000 88}

@@

```{r, correlation 5000 1010}

@@

```{r, correlation 5000 2020}

@@

```{r, boxplotting 5000 dif size}
  print("CA = mean_newt_pheno_By_mean_snake_pheno")
  print("CB = num_newts_By_num_snakes")
  print("CC = sum_newt_pheno_By_num_snake")
  print("CD = sum_snake_pheno_By_num_newt")

@@

print("Difference in amount of grids")
grid.arrange(G1, G4, G5, G6, nrow=2)

@@

```{r ga2 values}

@@

```{r, GA2 correlation}

@@

```{r, cor 5000 GA2 zooom}

@@

```{r, GA2 30000 correlation}

@@

```{r, cor 30000 GA2 mid zooom}

title="GA2 ZOOM Space: Newt VS Snake"
subt="30,000 Gen Grid 5 by 5"
x_label="Generation"
y_label="Correlation"
cors_by_gen(GA2_cor_30000_file[which(GA2_cor_30000_file$gen>=10000 & GA2_cor_30000_file$gen<11010),], title, subt, x_label, y_label)

title="ZOOM Phenotpe by Gen For all Sims"
subt="GA2 30,000 Gen"
y_label="Phenotype"
trends_by_gen(GA2_GA_30000_file[ which(GA2_GA_30000_file$gen>=10000 & GA2_GA_30000_file$gen<11010),], title, subt, x_label, y_label)

@@

>>>>

```{r, Cor data gather, cache=TRUE}
#Reading in the names of files to cycle through them
lit_5000_files <- list.files(path="~/Desktop/cor_5000/cor_byUnit/text_files", pattern="GA5_lit_", full.names=TRUE, recursive=FALSE)
cor_5000_files <- list.files(path="~/Desktop/cor_5000/cor_byUnit/text_files", pattern="GA5_cor_", full.names=TRUE, recursive=FALSE)

@@

```{r, reusable figures, results='hide', message=FALSE, warning=FALSE, cache=TRUE}

@@

boxplot_cors_by_type_gen <- function(file, subt, x_label, y_label, mycolor){

  plot <- ggplot(file) +
  ylim(-1.1, 1.1)+
  ggtitle(subt) +
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor("1"), y=as.numeric(C1)), color = mycolor)+
  geom_boxplot(aes(x=factor("2"), y=as.numeric(C2)), color = mycolor)+
  geom_boxplot(aes(x=factor("3"), y=as.numeric(C3)), color = mycolor)+
  geom_boxplot(aes(x=factor("4"), y=as.numeric(C4)), color = mycolor)+
  theme(plot.title = element_text(size=5)) +
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

@@

```{r, lest see that correlation, fig.height = 8, fig.width = 10, cache=TRUE}

@@

```{r, cor 5000 zooom, fig.height = 8, fig.width = 10, cache=TRUE}

@@

```{r, correlation 20000, fig.height = 8, fig.width = 10, cache=TRUE}

@@

```{r, correlation 100000, fig.height = 8, fig.width = 10, cache=TRUE}

@@

```{r, boxplotting 5000, fig.height = 8, fig.width = 8, cache=TRUE}
  print("CA = mean_newt_pheno_By_mean_snake_pheno")
  print("CB = num_newts_By_num_snakes")
  print("CC = sum_newt_pheno_By_num_snake")
  print("CD = sum_snake_pheno_By_num_newt")

@@

grid.arrange(G1, G2, G3, nrow=2, top=textGrob("Last 1,000 Gens for Different Simulation Lengths",gp=gpar(fontsize=16,font=3)))

@@

```{r, correlation 5000 88, fig.height = 8, fig.width = 10, cache=TRUE}

@@

```{r, correlation 5000 1010, fig.height = 8, fig.width = 10, cache=TRUE}

@@

```{r, correlation 5000 2020, fig.height = 8, fig.width = 10, cache=TRUE}

@@

```{r, boxplotting 5000 dif size, fig.height = 8, fig.width = 8, cache=TRUE}
  print("CA = mean_newt_pheno_By_mean_snake_pheno")
  print("CB = num_newts_By_num_snakes")
  print("CC = sum_newt_pheno_By_num_snake")
  print("CD = sum_snake_pheno_By_num_newt")

@@

grid.arrange(G1, G4, G5, G6, nrow=2, top=textGrob("1000 Gen at Differnt Time Points",gp=gpar(fontsize=16,font=3)))

@@

```{r ga2 values, cache=TRUE}

@@

The next section contains the results of the correlation calculations and mean phenotypes in the GA2 experiment. The first figure shows 4 correlation results, which were collected with a 5 by 5 grid. The correlation between mean newt phenotype and mean snake phenotype is the red line (CA). The correlation between the number of newts and the number of snakes is the green line (CB). The correlation between the sum of toxicity and the number of snakes is the blue line (CC). The correlation between the amount of resistance and the number of newt is the purple line (CD). This figure shows all 25 simulations in this set. The second figure shows the mean newt phenotype (red) the snake mean phenotype (blue) and the difference between the snake mean phenotype and the newt mean phenotype (black) for all 25 simulations.

```{r, GA2 correlation, fig.height = 8, fig.width = 10, cache=TRUE}

@@

The first figure shows the 4 correlation calculations collected every 10 generations for a 5,000 simulation. The first thing that I noticed in the first figure was a diminished CA correlation value. I did notice that CA values got a bit higher when the mean newt and snake values on the second plot intersected. These results were similar to the results seen in the GA1 experiment, with the exception of lower CA values. CB, CC, CD are all negative and generally fluctuate between 0 and -0.5. CA has a larger positive fluctuation between 0.0 and 1. The second figure shows the interaction of the mean snake and mean newt phenotypes. In this figure the phenotypes are often close to each other. I looked at both figures and wondered if maybe 5,000 generations was not enough time for a co-evolutionary relationship to form or stabilize. I also wondered why CA was lower (and did not start at 1) in this set of simulations.

#### Zooming in on Correlation Caculations for 5 by 5 grid GA2

Now, I am going to look at the zoomed in section of the GA2 5,000 generation simulations. I plotted the first 200 generations of my simulation and looked at the four correlation calculations and the newt & snake mean phenotype. Mostly, I want to look at where the correlations start off at and see how they immediately change. I want to compare the mean newt and snake phenotypes to the correlations.

```{r, cor 5000 GA2 zooom, fig.height = 8, fig.width = 10, cache=TRUE}

@@

The first thing I looked at was the second figure. The blue and red lines are not close together and it doesn't look like there a lot of movement in most of the simulations. Some of the red and blue lines decrease, and very few of the lines seem to increase. However, when looking at the correlation plots a lot of things are happening. There are some cases where things are falling and in other simulations things are going up. The blue, green, and purple generally start near 0. The red line usually starts off positive then decreases then goes back up. Its kind of weird how all four correlation lines can cluster at a certain time, but the mean phenotypes don't seem to change. I looked back at the zoomed in version of the GA1 experiment and found that the red line is a lot further away from the other lines in this experiment set. There was more of a trend in the GA1 experiment than in the GA2 experiment. Maybe it takes more time for co-evolution to occur when the variation between newts and snakes are the same.

## 30,000 Generation Results

In this section, I re-run the 25 GA2 simulations and increased the amount of generations to 30,000 (after looking at the 100,000 gen results). Here, I will again look at the four correlation calculations (CA, CB, CC, CD) and the mean newt and snake phenotypes. I hope to see more interaction between newt and snake mean phenotypes and a stabilization in the correlation calculations.

```{r, GA2 30000 correlation, fig.height = 8, fig.width = 10, cache=TRUE}

@@

After looking at the results, I noticed that the the red line (newt and snake phenotype correlation) generally stays high after a certain amount of generations. It is really difficult to see how the purple, blue, and green correlation lines interact. It seems like they are mostly overlap each other. It looks like these three correlations might become a bit more negative as the number of generations increase (at lease in some of the simulations). When looking at the second figure the mean newt and mean snake phenotype usually start out at with one species phenotype being higher and as the number of generations increase both phenotypes increase and eventually cross. There were four simulation where there was no crossing between mean newt and snake phenotype (B, F, H, Y). When snake mean phenotype was higher the correlation between newt and snake phenotype was diminished (F, H), but when newt mean phenotype was higher the correlation was a bit higher (B, Y).

#### Zooming in on 1,000 gen Period of Correlation Caculations for 5 by 5 grid

From the above section I wanted to zoom in on a 1,000 generation section where newt and snake phenotypes seem to be increasing. I choose to look at a section between 10,000 and 11,000 generations.

```{r, cor 30000 GA2 mid zooom, fig.height = 8, fig.width = 10, cache=TRUE}

title="GA2 midZOOM Space: Newt VS Snake"
subt="30,000 Gen Grid 5 by 5"
x_label="Generation"
y_label="Correlation"
cors_by_gen(GA2_cor_30000_file[which(GA2_cor_30000_file$gen>=10000 & GA2_cor_30000_file$gen<11010),], title, subt, x_label, y_label)

title="midZOOM Phenotpe by Gen For all Sims"
subt="GA2 30,000 Gen"
y_label="Phenotype"
trends_by_gen(GA2_GA_30000_file[ which(GA2_GA_30000_file$gen>=10000 & GA2_GA_30000_file$gen<11010),], title, subt, x_label, y_label)

@@

I mainly focus on looking for trends in the correlation results and the first one I noticed was how some of the four correlations seem to be following the same up and down pattern (most notably F, H, I, L, M). The blue, green and purple lines are mostly the same. The red line is usually positive and separated from the three other correlation lines. When I add in the second figure to what I observe in the first figure I start thinking about the how the difference in correlations might be dependent on which species currently has a higher phenotype (maybe over all or locally?). Potentially, when the newt mean phenotype is higher (overall) snakes in areas with high local newt phenotypes are under under selection to be more resistant. This selection might create a strong correlation between newt and snake phenotype, but when snakes meet or proceed the high newt phenotype the correlation gets weaker (S, U, V).

## Boxplot at Different Time Points

To get a better visual of what might be happening to the correlation calculation when I take them at different intervals. I decided to look at box plots. The first box plot I look at all four correlation calculations for each simulation separated by four 1,000 generation sections. The first section is the first 1,000 generation. The second section is taken from the 4,000- 5,000 generations. The third section is from the 10,000-11,000 generation and the last section is the lass 1,000 generation (29,000-30,000). As before, there are no labels on the x-axis (correlation calculation), the y-axis (correlation values), or the simulation letter (A starts on the top left and continues right and then to the next line). The four correlation calculations are printed below:

```{r, boxplotting 30000 dif time, fig.height = 8, fig.width = 8, cache=TRUE}
  print("CA = mean_newt_pheno_By_mean_snake_pheno")
  print("CB = num_newts_By_num_snakes")
  print("CC = sum_newt_pheno_By_num_snake")
  print("CD = sum_snake_pheno_By_num_newt")

title="Space: Newt VS Snake"
subt="25sims first 1000Gen, Grid 5 by 5"
x_label="Type"
y_label="Correlation"
G7<-boxplot_cors_by_gen(GA2_cor_30000_file[which(GA2_cor_30000_file$gen<=1000),], subt, x_label, y_label)

title="Space: Newt VS Snake"
subt="25sims 4000-5000 gen, Grid 5 by 5"
x_label="Type"
y_label="Correlation"
G8<-boxplot_cors_by_gen(GA2_cor_30000_file[which(GA2_cor_30000_file$gen>=4000 & GA2_cor_30000_file$gen<=5000),], subt, x_label, y_label)

title="Space: Newt VS Snake"
subt="25sims 10000-11000, Grid 5 by 5"
x_label="Type"
y_label="Correlation"
G9<-boxplot_cors_by_gen(GA2_cor_30000_file[which(GA2_cor_30000_file$gen>=10000 & GA2_cor_30000_file$gen<=11000),], subt, x_label, y_label)

title="Space: Newt VS Snake"
subt="25sims last 1000Gen 30000, Grid 5 by 5"
x_label="Type"
y_label="Correlation"
G10<-boxplot_cors_by_gen(GA2_cor_30000_file[which(GA2_cor_30000_file$gen>=29000),], subt, x_label, y_label)

grid.arrange(G7, G8, G9, G10, nrow=2, top=textGrob("1,000 Gen at Different Timepoints By Simulation",gp=gpar(fontsize=16,font=3)))

This figure is super difficult to read! The only thing I can tell is that the red box is consistently higher than the three other boxes. In the first 5,000 generations the purple box might be less negatively correlated in some of the simulations. However, for the most part the green, blue, and purple box plots look very similar. To see if the correlations were going up or down as the simulation kept running I decided to group the different box plots by which calculation I made.

The first graph (red) in the next figure is the correlation between newt and snake phenotype (CA) at the four different time points (1 = 0-1,000 gen, 2 = 4,000-5,000 gen, 3 = 10,000-11,000 gen, 4 = 29,000-30,000 gen). The second, third and fourth graph (green, blue, purple) is the correlation between number of newts and number of snakes (CB), the correlation between the amount of toxin and snakes (CC), and the correlation between resistance and number of newts (CD). Again, 1 = 0-1,000 gen, 2 = 4,000-5,000 gen, 3 = 10,000-11,000 gen, 4 = 29,000-30,000 gen.

```{r, regrouping the box plot, cache=TRUE}
# out.width = "1026", out.height = "528"
myvars <- c("gen", "mean_newt_pheno_By_mean_snake_pheno", "sim")
data_CA<-GA2_cor_30000_file[which(GA2_cor_30000_file$gen<=1000),][myvars]
names(data_CA)[names(data_CA) == 'mean_newt_pheno_By_mean_snake_pheno'] <- 'C1'
CA2<-GA2_cor_30000_file[which(GA2_cor_30000_file$gen>=4000 & GA2_cor_30000_file$gen<=5000),]$mean_newt_pheno_By_mean_snake_pheno
CA3<-GA2_cor_30000_file[which(GA2_cor_30000_file$gen>=10000 & GA2_cor_30000_file$gen<=11000),]$mean_newt_pheno_By_mean_snake_pheno
CA4<-GA2_cor_30000_file[which(GA2_cor_30000_file$gen>=29000),]$mean_newt_pheno_By_mean_snake_pheno
data_CA$C2 <- CA2
data_CA$C3 <- CA3
data_CA$C4 <- CA4

myvars <- c("gen", "num_newts_By_num_snakes", "sim")
data_CB<-GA2_cor_30000_file[which(GA2_cor_30000_file$gen<=1000),][myvars]
names(data_CB)[names(data_CB) == 'num_newts_By_num_snakes'] <- 'C1'
CB2<-GA2_cor_30000_file[which(GA2_cor_30000_file$gen>=4000 & GA2_cor_30000_file$gen<=5000),]$num_newts_By_num_snakes
CB3<-GA2_cor_30000_file[which(GA2_cor_30000_file$gen>=10000 & GA2_cor_30000_file$gen<=11000),]$num_newts_By_num_snakes
CB4<-GA2_cor_30000_file[which(GA2_cor_30000_file$gen>=29000),]$num_newts_By_num_snakes
data_CB$C2 <- CB2
data_CB$C3 <- CB3
data_CB$C4 <- CB4

myvars <- c("gen", "sum_newt_pheno_By_num_snake", "sim")
data_CC<-GA2_cor_30000_file[which(GA2_cor_30000_file$gen<=1000),][myvars]
names(data_CC)[names(data_CC) == 'sum_newt_pheno_By_num_snake'] <- 'C1'
CC2<-GA2_cor_30000_file[which(GA2_cor_30000_file$gen>=4000 & GA2_cor_30000_file$gen<=5000),]$sum_newt_pheno_By_num_snake
CC3<-GA2_cor_30000_file[which(GA2_cor_30000_file$gen>=10000 & GA2_cor_30000_file$gen<=11000),]$sum_newt_pheno_By_num_snake
CC4<-GA2_cor_30000_file[which(GA2_cor_30000_file$gen>=29000),]$sum_newt_pheno_By_num_snake
data_CC$C2 <- CC2
data_CC$C3 <- CC3
data_CC$C4 <- CC4

myvars <- c("gen", "sum_snake_pheno_By_num_newt", "sim")
data_CD<-GA2_cor_30000_file[which(GA2_cor_30000_file$gen<=1000),][myvars]
names(data_CD)[names(data_CD) == 'sum_snake_pheno_By_num_newt'] <- 'C1'
CD2<-GA2_cor_30000_file[which(GA2_cor_30000_file$gen>=4000 & GA2_cor_30000_file$gen<=5000),]$sum_snake_pheno_By_num_newt
CD3<-GA2_cor_30000_file[which(GA2_cor_30000_file$gen>=10000 & GA2_cor_30000_file$gen<=11000),]$sum_snake_pheno_By_num_newt
CD4<-GA2_cor_30000_file[which(GA2_cor_30000_file$gen>=29000),]$sum_snake_pheno_By_num_newt
data_CD$C2 <- CD2
data_CD$C3 <- CD3
data_CD$C4 <- CD4

subt="CAs"
x_label="Type"
y_label="Correlation"
CAs<-boxplot_cors_by_type_gen(data_CA, subt, x_label, y_label, mycolor="#F8766D")

subt="CBs"
x_label="Type"
y_label="Correlation"
CBs<-boxplot_cors_by_type_gen(data_CB, subt, x_label, y_label, mycolor="#7CAE00")

subt="CCs"
x_label="Type"
y_label="Correlation"
CCs<-boxplot_cors_by_type_gen(data_CC, subt, x_label, y_label, mycolor="#00BFC4")

subt="CDs"
x_label="Type"
y_label="Correlation"
CDs<-boxplot_cors_by_type_gen(data_CD, subt, x_label, y_label, mycolor="#C77CFF")

```{r, regrouping the box plot figure, fig.height = 8, fig.width = 8, cache=TRUE}

grid.arrange(CAs, CBs, CCs, CDs, nrow=2, top = textGrob("1,000 Gens Grouped by Caculation",gp=gpar(fontsize=16,font=3)))

The trend between time points was easier to see in this figure. Generally, the correlation between newt and snake phenotype increased at the number of generation increased (except F). The correlation between number of newts and number of snakes (CB), the correlation between the amount of toxin and snakes (CC), and the correlation between resistance and number of newts (CD) generally stay about the same. There are instances where they seem to move in a curved motion (like ~). It makes me wonder what makes these correlations stay between 0 and -0.5. I am also curious to how the correlations are connect to each other, if they are connect to each other.

## Future Ideas

- How does interaction rate effect the correlation calculations?
- How does avoidance change correlation?
- When I run multiple trials how do I show that information?

## Conclusion

Write after meeting with Peter
